Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-05 android-stacks.tex
intro

\begin{abstract}
The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptinal conditions 
(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
fundamental part of such apps. However, studies have shown that such mechanisms are the less 
understood and tested parts of the system. As a consequence they may affect the system the 
other way around, making the system even less robust, leading to failures such as uncaught exceptions 
and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
This paper reports a study which mined the stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Google Code (157 projs.). 
The goal of this study was to: investigate the whether the exception stack traces
 can reveal common characteristics on the exception-related code that
may be favoring the introduction of the aforementioned failures, which would be responsible for 
decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some fragilities consistently detected in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by JNI interface;
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
revealing that Java hybrid exception model is not fully used according to its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
as the root cause of an exception stack trace.
This study provide important insights on how such fragilities impose challenges 
to application robustness not only in Android environment but also to other 
systems based on Java exception model.

@@

Along the recent years, we have witnessed an astonishing increase in the number 
mobile applications. If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face an increasing number of threats to application robustness
 arising from: faults on underlying middleware and hardware (e.g., GPS receiver, camera, sensors);
faults on third party services and libraries; compatibility issues ~\cite{McDon13}; 
 memory and battery restrictions~\cite{Zhang12}; concurrency issues ~\cite{ama2012}; among others.

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps. The exception handling mechanism ~\cite{goodenough1975exception},
embedded in many mainstream programming languages, such as Java, C++ and C\#,
 is one of the most used techniques for detecting and recovering from such exceptional conditions. 
Android, the leadling plaform for mobile applications~\cite{gartner} (thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
 own device)  reuses the embedded Java Exception model to detect and handle 
 exceptional conditions.

Studies have shown, however, that the exception-related code is generally poorly understood and the
 less tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
As a consequence they may affect the system the other way around.
The exception-related code may favor the introduction of failures such as 
uncaught exceptions~\cite{jo2004uncaught, zhang2012amplifying}
which make the system even less robust~\cite{coelho2011unveiling}.

@@

Until recent years, the availability of data, reporting such failures, was scarce. 
It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their version control systems as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Google Code.
Developers and users of open-source projects hosted these sites 
constantly uses the provided infra-strucute to share information across 
the software development (including exception stack traces detected).

@@

This study performs a post mortem analysis of the exception stack traces embedded on issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study was to: investigate the whether the reported exception stack traces
 can reveal common fragilities on the exception-related code which may be favoring 
the introduction of failures such as uncaught exceptions - and as a consequence
 contributing to deacrease the robustness of apps.
To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented with additional information
extracted from JDK, Android Plaform and the applications using byte code 
and source code analysis. Manual inspection was used to leverage the understanding 
of exception stack traces and support further discussions and insights.

@@

   \item  Unexpected exception wrappings (e.g., Error wrapping checked and runtime
    exceptions and vice-versa), indicating that the purpose of Java's hybrid exception model may not have been adequately used.

   \item Undocumented runtime exceptions signaled by third party code (i.e., libraries and Android platform). 
This study found 118 distinct methods defined on third party code and  reported as the root signalers of 
runtime exceptions. Only 1 of such methods (0,8\%) documented the explicitly thrown runtime exception 
in a Javadoc comment; and none of these methods included the exception as part of the exception 
interface (i.e., using throws clause).

   \item Undocumented checked exception signaled by native Android library. This finding points to a fragility 
on the way exceptions flow from JNI interface functions, which can lead to uncaught exceptions that are difficult do 
debug.

@@

@@

As we can see, although the study was based on a subset of Android projects, 
the findings are not limited to this setting. The detected fragilities make difficult
or sometimes infeasible for a Java developer to prevent failures such as 
uncaught exceptions (e.g., thrown silently by third-party utility code), and 
unintended handling ( e.g., when an Error is wrapped in a checked Exception).
Thereore, they impose obstacles not only to the developement of robust Android apps, 
but also to the developement of any robust Java-based system. 
Hence, the study results are relevant to: (i) Android and 
Java developers to familiarize common characteristics and fragilities of exception-related code 
which is the first step to help developers preventing and dealing with them; and (ii) designers of languages and tools 
to consider ways of reducing such fragilities in Android and Java environments.
The main contributions of this study is twofold:
\begin{itemize}

  \item  It performs a large scale post mortem analysis of exception stack traces reported on Android projects hosted in GitHub and Google Code.
  \item  It introduces ExceptionMiner, a tool developed to support exception stack trace mining and analysis.
    analysis.

% AVD: We have not looked at private issues, and the tool is described above.
%We strongly believe that, specially in the open-source environment,
%faults are not to be hidden in a private bug issue. Faults should be shared and
%discussed, so that a developer can learn from other projects mistakes.
%Currently, the search facilities of repositories are very limited, and the
%ExceptionMiner tool is a contribution in this direction.

@@

By convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application. The user-defined exceptions 
can be either checked or runtime. The Java specification~\cite{gosling2000java} 
suggests that user-defined exceptions should be checked (because doing so 
the callers of a method will know about the exceptions that a method can throw and so 
they can decide what to do about them). However, runtime exceptions can also be 
used to represent user-defined exceptional conditions.In fact, there is a long-lasting debate 
about the pros and cons of both approaches ~\cite{javatut,stackoverlow,debate},
\footnote{152 questions in Stackoverlow are related to this debate}.

@@

In Java, once an exception is thrown, the runtime environment looks for the nearest enclosing exception handler
(in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
 Looking for the handler on the invocation stack is claimed to increase software reusability, 
since the invoker of an operation can handle it in a wider context ~/cite{miller1997issues}.

@@

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review fragilities on the exception-related code -  
we opted for not restricting the analysis only on defect
issues\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}.
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue which are automatically generated. Therefore, we excluded such automatically generated
issues from our analysis.

@@

\emph{\textbf{Exception Type Analysis}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues.
The nidyke walks up the type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed the exception hierarchy of all exceptions defined on:
Android Platform (API level 19) - which includes all basic Java exceptions that 
can be thrown during the app execution. The exceptions thrown inside the apps
were classified using manual inspections as well as the exceptions thrown only
on previous versions of the Android platform.

@@

\subsection{Replication Package}
All the data used in this study will be publicly available at http://.....
Specifically we provide: (i) all issues related to Android projects found
in GITHub and Google Code used in this study; (iii) all stack traces extracted
from issues (ii) the results of manual inspection steps; (iv) the tool developed 
in this work to support stack trace extraction and distilling.

@@

\begin{table}
  \centering
  \begin{tabular}{lr}
    \hline
    \bfseries{Cause} &  \bfseries{\% Occurrences} \\
    \hline
      Programming Mistakes &  52,0\%\\ 
      Memory Restriction   &   4,1\% \\ 
      General Exceptions   &  6,7\%\\
    \hline
 \end{tabular}
  \caption{Identifying the concerns related to root exceptions}
  \label{tab:causes}
\end{table}

@@

We can see, from Table ~\ref{tab:typeroottab}, that contrating with the other origins, most of the 
 exceptions signaled on Android Libcore (i.e., the set of libraries reused by the Android) were 
checked exceptions. Such set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
org.apache.http, org.json, org.xml, javax. Signaling checked exceptions is considered a 
good practice (see best practice IV in Section~\ref{sec:best}) because by using checked exceptions a library can define a precise 
exception interface ~/cite{miller1997issues} to its clients.
 Since such libraries are heavily used  in several projects, such finding can be related to the library's maturity.

@@

To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
 signaled by  runtime environment (due to programming mistakes) since these exceptions 
should not be documented on the method signature. Then we inspected the code each method 
(defined either on Android or on libraries) signaling  runtime exceptions not-related to programming mistakes.
As a result of this inspection, we could observe that only 1 method out of 118 inspected methods
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

Such undocumented runtime exceptions represent a threat to system robustness, specially
when such exceptions are thrown then the third party code is invoked inside the application 
(e.g. libraries, or framework utility code). In such cases the client usually do not have access to 
the source code, and in the absence of
the exception documentation it is very difficult or even impossible for the client of such third party code to 
protect the application against undocumented runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

%SELECT repo, *
% FROM stacktracesummary where stack like '%getDeclaredMethods%';

@@

\subsection{What can the Exception Wrappings tell us?}

%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
%repo !='android'  AND ROOT_EXTENDS LIKE 'CHECKED%' AND EXCEPTION_EXTENDS LIKE 'RUNTIME'

@@

\subsection{On the need of rescues to the almost infeasible task of preventing uncaught exceptions}
In this study we could observe undocumented runtime exceptions thrown by third party code,
and even undocumented checked exception thrown by Java supporting library.
Such undocumented exceptions make difficult, and most of the times infeasible
for the client code to protect against `unforeseen" situations that may happen 
while calling a library code.

@@

\noindent\emph{Internal Validity} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases. Moreover, in this study we manually classified the exceptions 
related to programming mistakes. To ensure the quality of the analysis, 
we calculated the interrater agreement after three independent 
coders classified a randomly selected sample (of 25 exception 
types from the total 100); the interater agreement was high (96\%).

Another threat relates to the fact that parts of our analysis 
are based on the availability of stack traces on issues on Github and Googlecode. 
In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
representative of valid crash information of the applications. 
One way to mitigate such threat would be to access to the full 
set of crash data per application. Although some services exist 
to collect crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14},
they do not provide open access to the crash reports of their client applications.
In our study, we mitigated this threat by manually inspecting
the source code associated to a subset of reported exception stack traces.
This subset contained the stack traces related to the main findings 
of the study (e.g., `ùndocumented runtime and checked exceptions",
and cross-type wrappings"). Through the manual inspection
 we could map the stack traces information to real characteristics 
of the exception-related code of analyzed apps.

\noindent\emph{External Validity} First, we limited out analysis to open-source
Android apps. We are aware that the reported exceptions for commercial
apps can be different from the ones found in this study. However, 
in this study revealed that several reported exception stack traces 
were reporting root exceptions signaled by Android utility code and libraries. 
Moreover, since the ``fragilities" observed in this study are due to
characteristics of Java exception model, they impose challenges 
the robustness not ony to Android apps but also to other systems
 based on Java exception model.

Second, the results presented here were based on mining
issues on Github through the GHTorrent dataset. While comprehensive and
extensive, GHTorrent is not an exact replica of Github, so several issues might
be left out. Due to the way data collection works with GHTorrent, for projects
that are relatively inactive, GHTorrent might have not collected a significant
proportion of their issues. We did not investigate the extend of this threat on
our sample.

\footnote{to do: OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
FALAR QUE DESCOBERTAS FORAM DE LIBS JAVA. APESAR DO ESTUDO NAO SER FEITO EM PROJS JAVA.....
PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA}

\
\section{Related Work}
\label{sec:rele}

@@

\textit{Tools Support for Detecting Faults on Java Exception-Related Code.}

Manually analyzing the exception-related code and looking for defects, can easily 
become infeasible ~\cite{Robil00}. Hence, some tools have been proposed to detect faults 
on the exception handling code either statically (cabral, roberta,nelio,chang) 
or dynamicaly through testing (roberta). Both approaches, however, have inherent
 limitations. On the one hand, the static analysis tools needs to deal with the false 
positives and cannot adequately deal with dynamically loaded code (e.g. reflection  
[ref]). On the other hand the testing tools are llimited to the fact that the developer
 needs to create a test to exercise every exception, which is also costly.... (olhar tese). 
Moreover, both methods fail to account for exceptions implicitly signaled by the runtime 
environment (e.g. due to resource depletion which can be signaled in every statement invocation).

@@

In this paper we present an exploratory study in which we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Google Code. Overall it included 6,005 exception stack traces.
In this study, the information extracted from stack traces was used in combination 
with source code and bytecode analysis to identify common characteristics of exception 
stack traces and pin-point fragilities on the exception-related code. 
We call ``fragilities" characteristics on the exception-related code that favor the introduction
of  uncaught exceptions and unintended handling - which can contribute to 
degrade the robustness of applications.
Some fragilities consistently detected in this study were: 
(i) an abundance of NullPointerExceptions - it was mentioned as the root exception in approximately 
50\% of the analyzed projects. 
(ii) undocumented runtime exceptions thrown by third party code,
(iii) undocumented checked exceptions thrown by JNI native code,
(iv) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
As we can see, such fragilities can negatively affect not only the robustness of Android application 
but also the robustness of any Java-based application. 
Our results calls the attention for the need of tool and language support to help 
developers when dealing with such fragilities.

>>>>

\begin{abstract}
The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptional conditions 
(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
fundamental part of such apps. However, studies have shown that such mechanisms are the less 
understood and tested parts of the system. As a consequence they may affect the system the 
other way around, making the system even less robust, leading to failures such as uncaught exceptions 
and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
This paper reports a study which mined the stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Google Code (157 projs.). 
The goal of this study was to: investigate the whether the exception stack traces
 can reveal common characteristics on the exception-related code that
may be favoring the introduction of the aforementioned failures, which would be responsible for 
decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some fragilities consistently detected in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by a JNI interface function;
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
revealing that Java hybrid exception model is not fully used according to its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
as the root cause of an exception stack trace.
This study provides insights on such fragilities and the robustness threats  
they impose not only to Android apps but also to other systems 
based on Java exception model.

@@

Along the recent years, we have witnessed an astonishing increase in the number of
mobile applications. If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face an increasing number of threats to application robustness
 arising from: faults on underlying middleware and hardware (e.g., GPS receiver, camera, sensors);
faults on third party services and libraries; compatibility issues ~\cite{McDon13}; 
 memory and battery restrictions~\cite{Zhang12}; among others.

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps. The exception handling mechanism ~\cite{goodenough1975exception},
embedded in many mainstream programming languages, such as Java, C++ and C\#,
 is one of the most used techniques for detecting and recovering from such exceptional conditions. 
Android, the leading plaform for mobile applications~\cite{gartner} (thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
 own device)  reuses the embedded Java exception model to detect and handle 
 exceptional conditions.

Studies have shown, however, that the exception-related code is generally poorly understood and the
 less tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
As a consequence they may affect the system the other way around.
The exception-related code may favor the introduction of failures such as 
uncaught exceptions~\cite{jo2004uncaught, zhang2012amplifying} - 
which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.

@@

Until recent years, the availability of data, reporting such failures, was scarce. 
It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their version control systems as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Google Code.
Developers and users of open-source projects uses such provided
 infra-strucute to share information (e.g., exception stack traces~\cite{bettenburg2008makes,schroter2010stack}) across 
the software development.

@@

This study performs a post mortem analysis of the exception stack traces embedded on issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study was to: investigate whether the reported exception stack traces
 can reveal common fragilities on the exception-related code.  
We call ``fragilities" characteristics on the exception-related code 
that favor the introduction of failures such as uncaught 
exceptions -  contributing to decrease the applicaiton robustness.
To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented with additional information
extracted from JDK, Android Plaform and the applications using byte code 
and source code analysis. Manual inspection was used to leverage the understanding 
of exception stack traces and support further discussions and insights.

@@

@@

   \item  Cross-type exception wrappings (e.g., an OutOfMemoryError wrapped in a checked exception)
were detected on stack traces. Indicating that the purpose of Java's hybrid exception model may not have been adequately used.

   \item Undocumented runtime exceptions signaled by third-party code (i.e., libraries and Android platform). 
    In this study, none of the third-party methods explicitly signaling runtime exceptions included such exceptions
    on ``exception specifications" (i.e., a list of exceptions that a method might raise during its execution).

   \item Undocumented checked exception signaled by a  JNI interface function defined on Android Platform. 
    Some flows reported a checked exception thown by a JNI interface function, and  such exception was not  
    part of the ``exception interface" of JNI function nor the methods on its invokation chain - it can lead to 
   uncaught exceptions that are difficult to debug.

@@

Such outcomes point to threats not only to the developement of robust Android apps, 
but also to the developement of any robust Java-based system. For instance, in the absence of the 
``exception specification" of  a third-party code, it is difficult or 
even enfeasible for the developer to protect the code against ``unforeseen" exceptions. 
Since, in such cases the client usually do not have access to the source code, the undocumented exceptions may 
remain uncaught and lead to system crashes. Moreover, the cross-type exception wrappings 
(e.g., OutOfMemoryError is wrapped in a checked Exception) 
may lead to unintended handling problems - trying to handle an instance of OutOfMemoryError ``hiden" in 
a checked exception may lead the program to an unpredictable state.

Hence, the study results are relevant to Android and 
Java developers who may underestimate how likely are the
 exceptions related to programming mistakes, and how 
 tricky it can be to prevent the uncaught exceptions from third-party code
 in the absence of ``exception interfaces". Moreover, results are relevant to 
designers of languages and tools to consider ways of reducing such fragilities 
in Android and Java environments.

@@

By convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application. The user-defined exceptions 
can be either checked or runtime. The Java specification~\cite{gosling2000java} 
suggests that user-defined exceptions should be checked (because doing so 
the callers of a method will know about the exceptions that a method can throw and so 
they can decide what to do about them). However, runtime exceptions can also be 
used to represent user-defined exceptional conditions.In fact, there is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate},
\footnote{152 questions in Stackoverlow are related to this debate}.

@@

In Java, once an exception is thrown, the runtime environment looks for the nearest enclosing exception handler
(in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
 Looking for the handler on the invocation stack is claimed to increase software reusability, 
since the invoker of an operation can handle it in a wider context ~\cite{miller1997issues}.

@@

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review fragilities on the exception-related code -  
we opted for not restricting the analysis only on defect
issues\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{https://github.com/souzacoelho/exceptionminer}}.
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue which are automatically generated. Therefore, we excluded such automatically generated
issues from our analysis.

@@

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues.
The nidyke walks up the type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed the exception hierarchy of all exceptions defined on:
Android Platform (API level 19) - which includes all basic Java exceptions that 
can be thrown during the app execution. The exceptions thrown inside the apps
were classified using manual inspections as well as the exceptions thrown only
on previous versions of the Android platform.

@@

\subsection{Replication Package}
All the data used in this study will be publicly available at https://github.com/souzacoelho/exceptionminer
Specifically we provide: (i) all issues related to Android projects found
in GITHub and Google Code used in this study; (iii) all stack traces extracted
from issues (ii) the results of manual inspection steps; (iv) the tool developed 
in this work to support stack trace extraction and distilling.

@@

\begin{table}
  \centering
  \begin{tabular}{lr}
    \hline
    \bfseries{Cause} &  \bfseries{\% Occurrences} \\
    \hline
      Programming Mistakes &  51,96\%\\ 
      Memory Restriction   &   4,10\% \\ 
      General Exceptions   &  6,70\%\\
    \hline
 \end{tabular}
  \caption{Identifying the concerns related to root exceptions}
  \label{tab:causes}
\end{table}

@@

We can see, from Table ~\ref{tab:typeroottab}, that contrating with the other origins, most of the 
 exceptions signaled on Android Libcore (i.e., the set of libraries reused by the Android) were 
checked exceptions. Such set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
org.apache.http, org.json, org.xml, javax. Signaling checked exceptions is considered a 
good practice (see best practice IV in Section~\ref{sec:best}) because by using checked exceptions a library can define a precise 
exception interface ~\cite{miller1997issues} to its clients.
 Since such libraries are heavily used  in several projects, such finding can be related to the library's maturity.

@@

To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
 signaled by  runtime environment (due to programming mistakes) since these exceptions 
should not be documented on the method signature.

Then we inspected the code each method (defined either on Android  Aplication Framework or on libraries) 
explicitly signaling a runtime exception not-related to programming mistakes.
As a result of this inspection, we could observe that only 1 method out of 118 inspected methods
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

Such undocumented runtime exceptions represent a threat to system robustness, specially
when such exceptions are thrown then the third party code invoked inside the application 
(e.g. libraries, or framework utility code). In such cases the client usually do not have access to 
the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
 for the client to design the application to deal with ``unforeseen runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

@@

\subsection{What can the Exception Wrappings tell us?}
%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
%repo !='android'  AND ROOT_EXTENDS LIKE 'CHECKED%' AND EXCEPTION_EXTENDS LIKE 'RUNTIME'

@@

\subsection{On the need of rescues to the almost infeasible task of preventing uncaught exceptions}
In this study we could observe undocumented runtime exceptions thrown by third party code,
and even undocumented checked exception thrown by a JNI interface.
Such undocumented exceptions make difficult, and most of the times infeasible
for the client code to protect against `unforeseen" situations that may happen 
while calling a library code.

@@

\noindent\emph{Internal Validity.} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases. Moreover, in this study we manually classified the exceptions 
related to programming mistakes. To ensure the quality of the analysis, 
we calculated the interrater agreement after three independent 
coders classified a randomly selected sample (of 25 exception 
types from the total 100); the interater agreement was high (96\%).

Another threat relates to the fact that parts of our analysis 
are based on the availability of stack traces on issues on Github and Googlecode. 
In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
representative of valid crash information of the applications. 
One way to mitigate such threat would be to access to the full 
set of crash data per application. Although some services exist 
to collect crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14},
they do not provide open access to the crash reports of their client applications.
In our study, we mitigated this threat by manually inspecting
the source code associated to a subset of the reported exception stack traces.
Such subset comprise the stack traces related to the main findings 
of the study (e.g., `ùndocumented runtime and checked exceptions",
and cross-type wrappings").

\noindent\emph{External Validity.} We limited out analysis to a subset of
existing open-source Android apps. We are aware that the reported exceptions 
for commercial apps can be different from the ones found in this study, and that
this subset is a small percentage of existing apps.
Such threats to validity are similar to the ones imposed on the other empirical studies 
which also used free or open-source Android apps ~\cite{Linar13,McDon13,ahimed}.
Moreover,  several exception stack traces that support the findings of this study
refered to exceptions comming from methods defined on Android Application Framework
and third-party libraries.  Additionaly,  the ``fragilities" observed in this study are due to
characteristics of Java exception model, which can impose challenges 
the robustness of not ony to Android apps but also to other systems
 based on the same exception model.

\section{Related Work}
\label{sec:rele}

@@

\textit{Detecting Faults on Java Exception-Related Code.}  Some tools have been 
proposed to detect faults  
on the exception handling code either statically (cabral, roberta,nelio,chang) 
or dynamicaly through testing (roberta). Both approaches, however, have inherent
 limitations. On the one hand, the static analysis tools needs to deal with the false 
positives and cannot adequately deal with dynamically loaded code (e.g. reflection  
[ref]). On the other hand the testing tools are llimited to the fact that the developer
 needs to create a test to exercise every exception, which is also costly.... (olhar tese). 
Moreover, both methods fail to account for exceptions implicitly signaled by the runtime 
environment (e.g. due to resource depletion which can be signaled in every statement invocation).

@@

In this paper we present an exploratory study in which we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Google Code. Overall it included 6,005 exception stack traces.
In this study, the information extracted from stack traces was used in combination 
with source code and bytecode analysis to identify common characteristics of exception 
stack traces and pin-point fragilities on the exception-related code. 
We call ``fragilities" characteristics on the exception-related code that favor the introduction
of  uncaught exceptions and unintended handling - which can contribute to 
degrade the robustness of applications.
Some fragilities consistently detected in this study were: 
(i) an abundance of NullPointerExceptions - it was mentioned as the root exception in approximately 
50\% of the analyzed projects. 
(ii) undocumented runtime exceptions thrown by third party code,
(iii) undocumented checked exceptions thrown by JNI interface,
(iv) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
As we can see, such fragilities can negatively affect not only the robustness of Android application 
but also the robustness of any Java-based application. 
Our results calls the attention for the need of tool and language support to help 
developers when dealing with such fragilities.
